EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-09 03_edipack2.tex

\subsection{Lanczos based Diagonalization}\label{sSecHam}
In presence of symmetries (see \secu{sSecQNs}) the matrix representing
the Hamiltonian operator $\hat{H}$ takes a block diagonal form. Each block
contains the Hamiltonian matrix $H_{\SS_{\vec{Q}}}$ for a
given symmetry sector with quantum number $\vec{Q}$.
The analysis of the energy spectrum of the quantum impurity problems
thus reduces to the sequential solution of the sector secular problem
$$
H_{\SS_{\vec{Q}}}\ket{v}=E_{\SS_{\vec{Q}}}\ket{v}.
$$

Although the dimension of any such sector $\SS$ are much smaller than the
total Fock space dimension $D_\SS \ll D_\FF$ still the complete solution of
the related eigenvalue problem represents an unsurmountable problem
already with $N_\mathrm{s}\simeq 8$. 
In order to outwit this issue a number of
algorithms have been developed leveraging on the sparse nature of
$H_\SS$, e.g. Krylov or similar sub-space methods \cite{Lanczos,Arnoldi,Feast}.   
In \NAME the alghoritm of choice is P-ARPACK, which represents the
state-of-the-art Lanczos-based eigensolver with distributed memory
support~\cite{P-Arpack}. This method includes re-orthogonalization
within a block Lanczos algorithm which ensures convergence of the
requred eigenpairs. Using the input variable {\tt lanc\_method} it is
however possible to select a simple, but not as efficient, parallel
Lanczos algorithm.
The core of Krylov-based diagonalization algorithms is the Matrix-Vector Product
(MVP), i.e. the application of the Hamiltonian to a given input
vector:
$$
H_\SS \ket{v} \to \ket{w}\qquad \ket{v},\ket{w}\in \SS_{\vec{Q}}
$$

This operations easily takes over more than 80\% of the computation
time, so its optimization is a crucial aspect in any diagonalization
algorithm. In \onlinecite{Amaricci2022} we covered in full detail
the distrubuted memory massively parallel algorithms we implement 
to address this specific issue.
Here, using the same naming convention introduced in \onlinecite{Amaricci2022}, 
we briefly outline the main strategy for the implementation of the MVP
according to the value of {\tt ed\_mode}.

\paragraph{{\bf normal}}
This case relies on the tensor structure of the Fock space and the
symmetry sectors. In any given sector the electronic part of the
Hamiltonian matrix reads:
\begin{equation}
H_\SS = H_\mathrm{d}  + H_\up\otimes \mathbb{I}_\dw + \mathbb{I}_\up\otimes
H_\dw + H_\mathrm{nd}\,.
\label{HssNormal}
\end{equation}
where $H_d$ is a diagonal term containing the local part of the
Hamiltonian, there including the density-density terms of the
interaction,  $H_\sigma$ describes the hopping processes of the
electrons with spin $\sigma=\up,\dw$.
The term $H_{nd}$ contains all the remaining non-diagonal
elements which do not fit in the previous two components,
e.g. the spin-exchange and pair-hopping interaction terms.
Correspondingly, the any sector state $\ket{v}$ can be expressed in a matrix
basis $\hat{v}$ with rows (columns) running over spin $\up$ ($\down$)
configurations. Leveraging on this setup it is possible the MVP using the 
{\tt MPI\_All2AllV} which aims at optimizing locality in the memory
locality. Additional terms contained in $H_{nd}$ are treated using the
{\tt MPI\_AllGatherV} algorithm which includes a slight communication
overload ultimately reducing the parallel efficiency.

\paragraph{{\bf superc/nonsu2}}
In this two cases the symmetry enforce a specific relation between $\up$ and
$\dw$ configurations ultimately suppressing the benefits arising from
Fock space factorization.
The electronic part of the Hamiltonian matrix is usually organized in this structure:
$$
H_\SS = H_\mathrm{imp} + H_\mathrm{int} + H_\mathrm{bath} + H_\mathrm{imp-bath} 
$$
where, with almost obvious meaning of the labels, $H_\mathrm{imp}$ encodes
the non-interacting part of the impurity Hamiltonian determined by
$h^0_{\a\b\s\s'}$, $H_\mathrm{int}$ the interaction term, $H_\mathrm{bath}$ the
effective bath terms related to $h^p_{\a\b\s\s'}$ and finally
$H_\mathrm{imp-bath}$ describes the coupling between impurity and bath
proportional to $V^p_{\a\b\s\s'}$. 
In an MPI setup the first three terms contains elements local in the
memory of each node. These are stored separately from the non-local
ones in any {\tt sparse\_matrix} representing $H_\SS$. 
The MVP is carried out using the {\tt MPI\_AllGatherV} algorithm
which, as discussed in \onlinecite{Amaricci2022}, requires the
reconstruction of the distributed input array any node producing a
overflow in the MPI communication.

The diagonalization procedure is divided in two main steps: a global
setup and the actual diagonalization.
%
The setup step takes care of allocating the required memory,
setup the MPI environment and allocate the correct 
Matrix-Vector Product (MVP) procedure according to the symmetries of
the problem. 
This part is implemented in a set of independent modules  {\tt
  ED\_HAMILTONIAN\_?}, where {\tt ed\_mode=?}.
Specifically, any such module includes two functions.
The first, {\tt  build\_Hv\_sector\_?}, builds the required sector and
allocate the MVP function, i.e. point the abstract function {\tt
  sphtimesv\_p} to the correct procedure determined by the value of
{\tt ed\_sparse\_H=True/False}.
If {\tt True} the Hamiltonian terms in $H_\SS$
get evaluated and stored in a {\tt sparse\_matrix} instance, possibly
using automatic parallel storage, and then used to perform
the MVP. This part is implemented in the modules {\tt
  ED\_HAMILTONIAN\_?\_STORED\_HxV}
If {\tt ed\_sparse\_H} is {\tt False} the MVP is operated on-the-fly,
i.e. each element of $H_\SS$ is direcrtly applied to the given input
vector $\ket{v}$ to determine the outgoing vector $H_\SS\ket{v}$,
either in serial or parallel mode as implemented in {\tt
  ED\_HAMILTONIAN\_?\_DIRECT\_HxV}.

The function {\tt vecDim\_Hv\_sector\_?} returns the 
dimension of the vector to be used in the
MVP. In serial execution this number is just $D_\SS$ the dimension of
the sector, whereas in a parallel mode the returned value is the
dimension $d_i$ of the vector chunk per each node such that
$\sum_{i=1}^{N_\mathrm{nodes}}d_i = D_\SS$. Yet, the specific value of $d_i$
depends on the MPI algorithm used for the MVP function, which in turn
is determined by the symmetry of the problem, i.e. {\tt ed\_mode}.

The sector-by-sector diagonalization for each
value of {\tt ed\_mode} is implemented in the modules {\tt
  ED\_DIAG\_?} through the main functions {\tt
  diagonalize\_impurity\_?}. The diagonalization proceeds in three main
steps: i) setup the diagonalization by selecting all or a sub-set of the
sectors to analyze; ii) sequential diagonalization cycle over all the suitable symmetry
sectors; iii) analysis of the resulting {\tt state\_list} containins
all the eigensolutions to be conserved.

\subsection{Dynamical correlation functions}\label{sSecGF}
The determination of the low energy part of the Hamiltonian spectrum
enables the evaluation of dynamical correlation functions using Krylov
sub-space algorithm.
This is a key part of the library when using \NAME as an impurity
solver within DMFT framework.

Before entering into the implementation details we briefly outline the
generic algorithm. Let us consider the dynamical correlation function:
\begin{equation}
  \label{eqGaa}
  C_\AA = \ibra \TT_\pm[ \AA(t) \AA^+ ]\iket
  \end{equation}
where $\AA(t)=e^{iHt}\AA e^{-iHt}$, $\TT_\pm$ is the time-ordering operator for fermions or
bosons and $\ibra \AA \iket=\frac{1}{Z}\Tr{ \left[ e^{-\beta
      H}\AA \right]}$,  $Z=\sum_ne^{-\beta E_n}$, is the thermodynamic
average. Using spectral decomposition, the expression \equ{eqGaa} is reduced into Kallen-Lehman form:
\begin{equation}\label{KLgf}
  \begin{split}
    C_\AA(z) 
    &=  \ibra \AA \frac{1}{z-H} \AA^+\iket
    \mp \ibra \AA^+ \frac{1}{z+H} \AA\iket \cr
    %
    & =\frac{1}{Z}\sum_n e^{-\beta E_n}\sum_m
  \frac{ |\bra{\psi_m}\AA^+ \ket{\psi_n}|^2 }  {z-(E_m-E_n)}
  \mp
  \frac{ |\bra{\psi_m}\AA \ket{\psi_n}|^2 }  {z+(E_m-E_n)}\cr
\end{split}
\end{equation}
where $z\in\CCC$, 
$\ket{\psi_n}$, $E_n$ are the eigenpairs of the
Hamiltonian $H$ and $E_0$ is the groundstate energy.
While appealing, the previous expression requires the knowledge of the
entire Hamiltonian spectrum. A key simplification is obtained looking
at the first line in \equ{KLgf},
showing that the $C_\AA$ essentially corresponds to a given element of the inverse operator $(z-H)^{-1}$, i.e. the
resolvant. The Krylov sub-space method provides a controlled
approximation for this element, leveraging over the
reduction of the (sector) Hamiltonian to a partial tri-diagonal
form and returning an arbitrary number of amplitudes to (unknown)
excited states. To illustrate the algorithm let's consider the normalized initial state: 
$$
\ket{\phi_n}=\AA^+\ket{\psi_n}/\NN_n
$$
where  $\ket{\psi_n}\in\SS$ and $\NN_n=\sqrt{\bra{\psi_n}\AA
  \AA^+\ket{\psi_n}}$.
We can construct the Krylov basis $$\mathcal{K}_{N} (\ket{\phi_n})=\{\ket{\phi_n}, H\ket{\phi_n}, \dots,
H^N\ket{\phi_n}\}\equiv \{\ket{v^n_0}, \ket{v^n_1},\dots, \ket{v^n_N}
\}$$ with $1 \ll N \ll \DD_\SS$ by iteratively applying the (sector)
Hamiltonian through MVP functions. 
Any given eigenstate $\ket{\psi_n}$ has components along the Krylov basis $\mathcal{K}_{N} (\ket{\phi_n})$ which read: $\ket{\psi_n} =
\sum_i  \ibra v^n_i|\psi_n \iket  \ket{ v^n_i} =
\sum_i a^n_i \ket{v^n_i}$.   We can evaluate the operator components
$=\bra{\psi_m}\AA^+ \ket{\psi_n}\equiv \langle\psi_m| \phi_n\rangle =
\NN_n {a^{n*}_m}$. Inserting this expression into \equ{KLgf} we
obtain the following approximation to $C_\AA$:
\begin{equation}
  \label{eqGKrylov}
  \begin{split}
    C_\AA(z)  &\simeq \frac{1}{Z}\sum_n e^{-\beta E_n}
    \sum_{m=1}^{N} \frac{\bra{\psi_n}\AA \AA^+\ket{\psi_n} |a^n_m|^2}{
      z - (E_m-E_n)} \mp \frac{\bra{\psi_n}\AA^+ \AA\ket{\psi_n}
      |a^n_m|^2}{ z + (E_m-E_n)}\cr
    %
    &= \frac{1}{Z}\sum_n\sum_{m=1}^{N} \sum_{\nu=\pm}\frac{w^{\nu}_{mn}[\AA]}{z - dE^\nu_{mn}[\AA]} 
    = \frac{1}{Z}\sum_n
    \sum_{m=1}^{N} \sum_{\nu=\pm} g_\AA(\nu, w^{\nu}_{mn},  dE^\nu_{mn}) \cr
  \end{split}
\end{equation}
where we introduced the notation $g(\nu, w(\AA)^{\nu}_{mn},  dE(\AA)^\nu_{mn})$
for the {\tt gfmatrix} object containing the weights $w(\AA)^{\nu}_{mn}$ and
poles $dE(\AA)^{\nu}_{mn}$ for the operator $\AA$, for every initial state $\ket{\psi_n}$
contributing to the low energy spectrum, for every order $m$ of the
Krylov sub-space algorithm and for channel $\nu$.

A stringent limitation of this procedure is to get applied to diagonal
dynamical correlation functions of the form \equ{eqGaa}. However, as
we will show in the following, many practical application requires to
evaluate off-diagonal functions of the form:
$C_{\AA\BB}(z) = \ibra T_{\pm}[\AA(t) \BB^+]\iket$.
A direct extensin of the method is obtained considering suitable
auxiliary operators prepared with independent linear combination of
$\AA$ and $\BB$.
A notable example is to consider the two auxiiary operators $\OO=\AA +\BB$ and
$\PP=\AA-i\BB$. Using simple algebra it is then straightforward to
obtain the desired function $C_{\AA \BB}$ from the evaluation of
$C_\OO$ and $C_\PP$:
$$
C_{\AA\BB} = \frac{1}{2}\left[C_\OO + C_\PP - (1-i)C_\AA -(1-i)C_\BB\right]
$$

In \NAME the impurity Green's functions $G_{\a\b\s\s'}=\ibra \TT_\pm[ c_{\a\s}(t)
c_{\b\s'}^+ ]\iket$,
is contained in \texttt{ED\_GREENS\_NORMAL}. This wraps more
specific methods according to the symmetry of the problem expressed by
{\tt ed\_mode}. In the {\bf normal} mode also spin,
charge, pair and excitonic susceptibilities functions. 
Computationally, the construction of the Krylov basis
$\mathcal{K}_N(\OO\ket{\psi_n})$ for any eigenstate of the low energy
spectrum is the most time consuming step. As for the diagonalization
part, here a crucial speed-up is obtained by the parallel execution of
the MVP lying at the earth of the Hamiltonian tri-diaagonalization
process. The input variable {\tt lanc\_gfniter} controls the largest order of
the Krylov basis, i.e. the maximum number of excitations in
\equ{eqGKrylov}.  
Operationally, each symmetry mode {\tt ed\_mode=?} requires a different construction of
the Green's functions implemented in {\tt ED\_GF\_?}

\paragraph{{\bf normal}}
In this case the all the orbital dependent and spin-diagonal Green's
functions $G_{\a\b\s\s}$ needs to be evaluated. The orbital diagonal case
$\a=\b$ proceeds as discussed above considering $\AA=c_{\a\s}$. This
step makes use of the {\tt apply\_op\_C/CDG} functions in {\tt
  ED\_AUX\_FUNX} on any eigenstate $\ket{\psi_n}$ in the {\tt
  state\_list}, i.e. a global instance of {\tt sparse\_espace}
containing the low energy spectrum of the problem.
The sector Hamiltonian matrix in this mode is assumed to be real
symmetric. This brings a simplification in the evaluation of the
off-diagonal components $G_{\a\b\s\s}=G_{\b\a\s\s}$.
These components are thus evaluated considering the auxiliary operator
$\OO=(c_{\a\s} + c_{\b\s})$ and using the relation
$G_{\a\b\s\s}=\tfrac{1}{2}(C_\OO - G_{\a\a\s\s} - G_{\b\b\s\s})$.

The diagonal spin, charge and pair susceptibility terms
$\chi^{S^z}_{\a\a}$, $\chi^N_{\a\a}$ and $\chi^\Delta_{\a\a}$ are
evaluated employing, respectively, the operators $\AA=\sum_{\s\s'}c^+_{\a\s}\tau^z_{\s\s'}c_{\a\s'}\equiv
S^z_\a$, $\AA=\sum_{\s\s'}c^+_{\a\s}\tau^0_{\s\s'}c_{\a\s'}\equiv
N_\a$) and $\AA=c_{\a\dw}c_{\a\up}\equiv
\Delta_\a$, where
$\tau^{a=0,x,y,z}$ are the Pauli matrices.
Similarly, yhe off-diagonal terms are obtained by considering the operator $\AA=S^z_\a
+ S^z_\b$, $\AA=N_\a + N_\b$ or $\AA=\Delta_\a + \Delta_\b$. 
The excitonic susceptibility $\chi^T_{\a\b}$ is defined with respect
to the vector operator: $T^i_{\a\b} = \sum_{\s\s'}
c^+_{\a\s}\tau^i_{\s\s'}c_{\b\s'}$, respectively, the spin-singlet
($i=0$) and the spin-triplet excitons ($i=x,y,z$) operators.

\paragraph{{\bf superc}}
In the supercondutive case the Nambu orbital
dependent $s$-wave Green's function reads:
\begin{equation}
  \label{GFnambu}
  \hat{G}_{\a\b} =
  \begin{bmatrix}
    G_{\a\b\up\up} & F_{\a\b\up\dw} \\
    \bar{F}_{\a\b\dw\up} & \bar{G}_{\a\b\dw\dw}
  \end{bmatrix}  
\end{equation}
However, leveraging on the symmetries holding between the different
components we can reduce to evaluate only the top row components plus
auxiliary functions.
The calculation of the diagonal normal component $G_{\a\a\up\up}$ follows the
very same scheme of the {\bf normal} case. For the off-diagonal terms
however we do not rely on any symmetry relation but we consider 
two auxiliary functions $C_\OO$, $C_\PP$ with $\OO=c_{\a\up} +
c_{\b\up}$, $\PP=c_{\a\up} -ic_{\b\up}$ so that:
$G_{\a\b\up\up}=\tfrac{1}{2}[C_\OO + C_\PP -
(1-i)(G_{\a\a\up\up}+G_{\b\b\up\up})$.
The diagonal and off-diagonal anomalous terms $F_{\a\b\up\dw}$ need to
be evaluated using different combinations of creation/destruction
operators. First we evaluate the $\bar{G}_{\a\a\dw\dw}$ component as
an auxiliary term considering $\AA=c^+_{\a\dw}$. 
Next, we consider the two linear combinations:
$\TT = c_{\a\up} + c^+_{\b\dw}$, $\RR=c_{\a\up} -i c^+_{\b\dw}$
contributing, respectively, to the auxiliary functions $C_\TT$,
$C_\RR$. The searched anomalous functions reads:
$F_{\a\b\up\dw} = \tfrac{1}{2}[ C_\TT + C_\RR - (1-i)(G_{\a\a\up\up} +
\bar{G}_{\b\b\dw\dw})]$.

>>>>

\subsection{Lanczos based Diagonalization}\label{sSecHam}
In presence of symmetries (see \secu{sSecQNs}) the matrix representing
the Hamiltonian operator $\hat{H}$ can be reduced to a block-diagonal form. Each block corresponds to a symmetry sector
with a fixed set of quantum numbers $\vec{Q}$ and is represented by
a Hamiltonian matrix $H_{\SS_{\vec{Q}}}$. 
Consequently, the analysis of the quantum impurity energy spectrum reduces to solving
the secular equation within each individual symmetry sector:
$$
H_{\SS_{\vec{Q}}}\ket{v}=E_{\SS_{\vec{Q}}}\ket{v}.
$$

Even though the dimension of a given sector $\SS$ is much smaller than the
total Fock space dimension $D_\SS \ll D_\FF$, solving completely the
corresponding eigenvalue problem remains a formidable challenge
already for $N_\mathrm{s}\simeq 8$.
To outwit this, various
algorithms have been developed exploiting the sparse nature of
$H_\SS$, including Krylov subspace methods and related approaches
\cite{Lanczos,Arnoldi,Feast}.

In \NAME, the preferred algorithm is P-ARPACK, a state-of-the-art
Lanczos-based eigensolver with distributed memory support
\cite{P-Arpack}. This method incorporates re-orthogonalization within
a block Lanczos routine, ensuring the accurate convergence of the
desired eigenpairs. However, it is also possible to select a simpler,
though less efficient, parallel Lanczos algorithm using the input
variable {\tt lanc\_method}.

The core computational task in any Krylov-based diagonalization
algorithm is the Matrix-Vector Product (MVP), i.e., the application
of the Hamiltonian to a given input vector:
\begin{equation}
H_\SS \ket{v} \to \ket{w}\qquad \ket{v},\ket{w}\in \SS_{\vec{Q}}.
\end{equation}
This operation easily accounts for more than 80\% of the overall
computational effort, making its optimization a critical aspect of
high-performance diagonalization algorithms. In \onlinecite{Amaricci2022},
the distributed-memory, massively parallel strategies employed in \NAME
to optimize this step are discussed in full detail. Here, we provide
a brief outline of the main ideas, following the notation introduced
in \onlinecite{Amaricci2022}, and focusing on the impact of the choice
of {\tt ed\_mode}.

\paragraph{\bf normal}
In this case, the tensor structure of the Fock space and the sector
symmetries are fully leveraged. Within a given sector, the electronic
part of the Hamiltonian can be expressed as:
\begin{equation}
H_\SS = H_\mathrm{d} + H_\up\otimes \mathbb{I}\dw + \mathbb{I}\up\otimes
H_\dw + H_\mathrm{nd},
\label{HssNormal}
\end{equation}
where $H_\mathrm{d}$ represents the diagonal (local) part of the
Hamiltonian, including density-density interactions, while
$H_\sigma$ captures hopping processes for electrons with spin
$\sigma=\up,\dw$. The term $H_\mathrm{nd}$ encompasses all remaining
non-diagonal contributions, such as spin-exchange and pair-hopping
terms.

In this formulation, a sector state $\ket{v}$ can be
represented in a matrix basis $\hat{v}$ with rows (columns)
corresponding to $\up$ ($\dw$) configurations. This structure allows
for efficient MVP operations using {\tt MPI\_All2AllV}, which aims to
optimize memory locality. However, terms in $H_\mathrm{nd}$ that
break this structure require {\tt MPI\_AllGatherV}, introducing a
slight communication overhead and reducing parallel efficiency.

\paragraph{\bf superc/nonsu2}
For these modes, symmetries enforce specific constraints between
$\up$ and $\dw$ configurations, preventing the direct use of the
factorized Fock space representation. The electronic
part of the Hamiltonian is typically structured as:
\begin{equation}
H_\SS = H_\mathrm{imp} + H_\mathrm{int} + H_\mathrm{bath} + H_\mathrm{imp-bath},
\end{equation}
where $H_\mathrm{imp}$ encodes the local impurity Hamiltonian
determined by $h^0_{\a\b\s\s'}$,
$H_\mathrm{int}$ contains the interaction terms, $H_\mathrm{bath}$
describes the effective bath contributions corresponding to
$h^p_{\a\b\s\s'}$, and $H_\mathrm{imp-bath}$ accounts for the
impurity-bath couplings $V^p_{\a\b\s\s'}$.
In an MPI setup, the first three terms contains also elements local to each node and
stored separately within each {\tt sparse\_matrix} instance representing $H_\SS$.
The MVP is carried out using the {\tt MPI\_AllGatherV} algorithm which, as noted in
\onlinecite{Amaricci2022}, requires reconstruction of distributed
arrays and incurs significant communication overhead.

The diagonalization process itself is divided into two primary phases:

\paragraph{\bf Global Setup}. This phase allocates the required memory, initializes
the MPI environment, and configures the appropriate MVP procedure
based on the symmetries of the problem. This logic is implemented in
a set of independent modules, {\tt ED\_HAMILTONIAN\_?}, where
{\tt ed\_mode=?}. Any module includes two functions: {\tt build\_Hv\_sector\_?} and  {\tt vecDim\_Hv\_sector\_?}.

The {\tt build\_Hv\_sector\_?} function builds the symmetry sector and
allocates the MVP function but setting the shared abstract function pointer {\tt
  sphtimesv\_p}. This latter step is controlled by the value of
{\tt ed\_sparse\_H=True/False}. If {\tt True} the Hamiltonian $H_\SS$
gets stored in a {\tt sparse\_matrix} instance and used to perform
the MVPs. The corresponding algorithms are implemented in the modules {\tt
  ED\_HAMILTONIAN\_?\_STORED\_HxV}
If {\tt ed\_sparse\_H} is {\tt False} the MVP is operated on-the-fly,
i.e. each element of $H_\SS$ is directly applied to the input
vector $\ket{v}$, either in serial or parallel mode as implemented in {\tt
  ED\_HAMILTONIAN\_?\_DIRECT\_HxV}.

The function {\tt vecDim\_Hv\_sector\_?} returns the 
dimension of the vector used in the
MVP. In a MPI-parallel execution the returned value is the
dimension $d_i$ of the vector chunk per each node such that
$\sum_{i=1}^{N_\mathrm{nodes}}d_i = D_\SS$. Yet, the specific value of $d_i$
depends on the MPI algorithm used for the MVP function.

\paragraph{\bf Diagonalization}
This step is managed by the {\tt ED\_DIAG\_?}
modules, which contain the main functions {\tt diagonalize\_impurity\_?}.
This phase includes: (i) selecting sectors for diagonalization, (ii)
performing diagonalization within each sector, and (iii) analyzing
the resulting {\tt state\_list} of conserved eigenstates.

\subsection{Dynamical correlation functions}\label{sSecGF}
The determination of the low energy part of the Hamiltonian spectrum
enables the evaluation of Dynamical Correlation Functions (DCF) using Krylov
sub-space algorithm.
This capability is central to the library when using \NAME as an
impurity solver within the DMFT framework.

Before delving into the implementation specifics, we outline the
generic approach. Consider the generic DCF:
\begin{equation}
  \label{eqGaa}
  C_\AA = \ibra \TT_\pm[ \AA(t) \AA^+ ]\iket
\end{equation}
where $\AA(t)=e^{iHt}\AA e^{-iHt}$, $\TT_\pm$ is the time-ordering
operator for fermions ($-$) or
bosons ($-$), and $\ibra \AA \iket=\frac{1}{Z}\Tr{ \left[ e^{-\beta
      H}\AA \right]}$, with $Z=\sum_ne^{-\beta E_n}$, the thermodynamic
average.
Using spectral decomposition, the expression \equ{eqGaa} reduces to a
K\"all\`en-Lehmann form:
\begin{equation}\label{KLgf}
  \begin{split}
    C_\AA(z) 
    &=  \ibra \AA \frac{1}{z-H} \AA^+\iket
    \mp \ibra \AA^+ \frac{1}{z+H} \AA\iket \cr
    %
    & =\frac{1}{Z}\sum_n e^{-\beta E_n}\sum_m
  \frac{ |\bra{\psi_m}\AA^+ \ket{\psi_n}|^2 }  {z-(E_m-E_n)}
  \mp
  \frac{ |\bra{\psi_m}\AA \ket{\psi_n}|^2 }  {z+(E_m-E_n)}\cr
\end{split}
\end{equation}
where $z\in\CCC$,  $\ket{\psi_n}$, $E_n$ are the eigensolutions of the
Hamiltonian $H$. 
This form is appealing but computationally prohibitive, as it requires
the full Hamiltonian spectrum. However, the first line in \equ{KLgf}
highlights that the $C_\AA$ essentially corresponds to a specific
matrix element of the resolvant operator $(z-H)^{-1}$, which can be
efficiently approximated using the Krylov sub-space method.

To illustrate this approach, consider the normalized initial state: 
$$
\ket{\phi_n}=\AA^+\ket{\psi_n}/\NN_n
$$
where  $\ket{\psi_n}\in\SS$ and $\NN_n=\sqrt{\bra{\psi_n}\AA
  \AA^+\ket{\psi_n}}$.
The Krylov basis is then constructed as $$\mathcal{K}_{N} (\ket{\phi_n})=\{\ket{\phi_n}, H\ket{\phi_n}, \dots,
H^N\ket{\phi_n}\}\equiv \{\ket{v^n_0}, \ket{v^n_1},\dots, \ket{v^n_N}
\}$$ with $1 \ll N \ll \DD_\SS$, obtainaed by repeated applications of
the (sector) Hamiltonian via MVP functions.

Given that any eigenstate $\ket{\psi_n}$ can be expressed within the
Krylov basis as
$\ket{\psi_n} = \sum_i  \ibra v^n_i|\psi_n \iket  \ket{ v^n_i} =
\sum_i a^n_i \ket{v^n_i}$. we can approximate the expression
\equ{KLgf} as:  
\begin{equation}
  \label{eqGKrylov}
  \begin{split}
    C_\AA(z)  &\simeq \frac{1}{Z}\sum_n e^{-\beta E_n}
    \sum_{m=1}^{N} \frac{\bra{\psi_n}\AA \AA^+\ket{\psi_n} |a^n_m|^2}{
      z - (E_m-E_n)} \mp \frac{\bra{\psi_n}\AA^+ \AA\ket{\psi_n}
      |a^n_m|^2}{ z + (E_m-E_n)}\cr
    %
    &= \frac{1}{Z}\sum_n\sum_{m=1}^{N} \sum_{\nu=\pm}\frac{w^{\nu}_{mn}[\AA]}{z - dE^\nu_{mn}[\AA]} 
    = \frac{1}{Z}\sum_n
    \sum_{m=1}^{N} \sum_{\nu=\pm} g_\AA(\nu, w^{\nu}_{mn},  dE^\nu_{mn}) \cr
  \end{split}
\end{equation}
where the terms are grouped as a sum over spectral weights and
poles. We introduced the notation $g(\nu, w(\AA)^{\nu}_{mn},  dE(\AA)^\nu_{mn})$
for the {\tt gfmatrix} object containing the weights $w(\AA)^{\nu}_{mn}$ and
poles $dE(\AA)^{\nu}_{mn}$ for the operator $\AA$, for every initial state $\ket{\psi_n}$
contributing to the low energy spectrum, for every order $m$ of the
Krylov sub-space algorithm and for channel $\nu$.
This approach significantly reduces the computational burden, as it
relies on a much smaller set of basis states.

However, this method is typically limited to diagonal DCF like
\equ{eqGaa}. In practice, many applications requires also to build off-diagonal
functions of the form:
$C_{\AA\BB}(z) = \ibra T_{\pm}[\AA(t) \BB^+]\iket$.
This can addressed by introducing auxiliary operators, such as $\OO=\AA +\BB$ and
$\PP=\AA-i\BB$, enabling the extraction of the desired function via
simple algebraic combinations:
$$
C_{\AA\BB} = \frac{1}{2}\left[C_\OO + C_\PP - (1-i)C_\AA -(1-i)C_\BB\right]
$$

In \NAME, the computation of impurity Green's functions, such as
$G_{\a\b\s\s'}=\ibra \TT_\pm[ c_{\a\s}(t) c_{\b\s'}^+ ]\iket$,
is handled by the \texttt{ED\_GREENS\_FUNCTIONS} module.
This
module integrates more specialized methods based on the symmetry
classification defined by the \texttt{ed\_mode} parameter.
Moreove, in the {\bf normal} mod, it also includes spin, charge, pair, and excitonic
susceptibility functions, providing a comprehensive framework for
dynamical response calculations.
From a computational perspective, the construction of the Krylov basis
$\mathcal{K}_N(\OO\ket{\psi_n})$ for each eigenstate within the low-energy
spectrum is typically the most resource-intensive step.
As with the
diagonalization process, a significant performance gain is achieved through
the parallel execution of the Matrix-Vector Product (MVP) at the core of
the Hamiltonian tri-diagonalization algorithm.
The input variable
\texttt{lanc\_gfniter} regulates the maximum order of the Krylov basis,
determining the upper limit on the number of excitations considered in
\equ{eqGKrylov}.
Operationally, each symmetry mode (\texttt{ed\_mode=?}) requires a distinct
strategy for Green's function construction, implemented in the corresponding
\texttt{ED\_GF\_?} modules.

\paragraph{{\bf normal}}
In this mode, all orbital-dependent and spin-diagonal Green's 
functions, $G_{\a\b\s\s}$, must be evaluated. The diagonal 
case, where $\a=\b$, is straightforward and involves applying 
the operator $\AA=c_{\a\s}$ to any eigenstate $\ket{\psi_n}$ 
from the global {\tt state\_list} of the low-energy spectrum, 
which is represented by the {\tt sparse\_espace} object. This 
operation relies on the functions {\tt apply\_op\_C/CDG} in 
{\tt ED\_AUX\_FUNX}.

The sector Hamiltonian matrix in this mode is typically assumed 
to be real symmetric, which significantly simplifies the evaluation 
of the off-diagonal terms. In this case, the relation 
$G_{\a\b\s\s}=G_{\b\a\s\s}$ holds, allowing the computation of 
off-diagonal components using an auxiliary operator 
$\OO = c_{\a\s} + c_{\b\s}$ and the identity:
$G_{\a\b\s\s}=\tfrac{1}{2}(C_\OO - G_{\a\a\s\s} - G_{\b\b\s\s})$.

The diagonal spin, charge and pair susceptibility terms
$\chi^{S^z}_{\a\a}$, $\chi^N_{\a\a}$ and $\chi^\Delta_{\a\a}$
are constructed using the operators $\AA = \sum_{\s\s'} c^+_{\a\s}
\tau^z_{\s\s'} c_{\a\s'} \equiv S^z_\a, $, 
$\AA = \sum_{\s\s'} c^+_{\a\s} \tau^0_{\s\s'} c_{\a\s'} \equiv N_\a, $
and $\AA = c_{\a\dw}c_{\a\up} \equiv \Delta_\a$,
where $\tau^{a=0,x,y,z}$ are the Pauli matrices.
Off-diagonal terms are similarly evaluated using the operators
$\AA=S^z_\a + S^z_\b$, $\AA=N_\a + N_\b$ or $\AA=\Delta_\a + \Delta_\b$. 
Excitonic susceptibilities, $\chi^T_{\a\b}$, are defined with 
respect to the vector operator: $T^i_{\a\b} = \sum_{\s\s'}
c^+_{\a\s}\tau^i_{\s\s'}c_{\b\s'}$, where $i=0$ represents the
spin-singlet exciton, and $i=x,y,z$  correspond to the spin-triplet excitons.

\paragraph{{\bf superc}}
In the supercondutive case, the Nambu orbital-dependent $s$-wave
Green's function reads:
\begin{equation}
  \label{GFnambu}
  \hat{G}_{\a\b} =
  \begin{bmatrix}
    G_{\a\b\up\up} & F_{\a\b\up\dw} \\
    \bar{F}_{\a\b\dw\up} & \bar{G}_{\a\b\dw\dw}
  \end{bmatrix}  
\end{equation}
Exploiting symmetries between matrix components, it is 
sufficient to compute only the top row elements along with 
a few auxiliary functions.

The diagonal normal component $G_{\a\a\up\up}$ can be 
evaluated using the same approach as in the {\bf normal} 
case. However, for the off-diagonal terms $G_{\a\b\up\up}$, 
we forgo symmetry arguments and instead define two 
auxiliary operators, $\OO = c_{\a\up} + c_{\b\up}$ and 
$\PP = c_{\a\up} - i c_{\b\up}$. This allows us to express 
the Green's function as::
$G_{\a\b\up\up}=\tfrac{1}{2}[C_\OO + C_\PP -
(1-i)(G_{\a\a\up\up}+G_{\b\b\up\up})$.

Evaluating the diagonal and off-diagonal anomalous terms 
$F_{\a\b\up\dw}$ requires distinct combinations of 
creation and annihilation operators. First, the component 
$\bar{G}_{\a\a\dw\dw}$ is evaluated as an auxiliary term 
using $\AA = c^+_{\a\dw}$. Then, we construct two further 
linear combinations:
\begin{align*}
\TT &= c_{\a\up} + c^+_{\b\dw}, \\
\RR &= c_{\a\up} - i c^+_{\b\dw},
\end{align*}
which contribute to the auxiliary functions $C_\TT$ and 
$C_\RR$, respectively. The final expression for the 
anomalous function reads:
\begin{equation}
F_{\a\b\up\dw} = \tfrac{1}{2} \left[ C_\TT + C_\RR - 
(1-i)(G_{\a\a\up\up} + \bar{G}_{\b\b\dw\dw}) \right].
\end{equation}

\paragraph{{\bf nonsu2}}
In the {\bf nonsu2} case, all Green's function components 
must be explicitly evaluated, as the spin and orbital 
symmetries are not conserved. For the diagonal terms 
$G_{\a\a\s\s}$, the procedure follows the same approach 
outlined in the {\bf normal} case.
